---
title: "445-13"
output: pdf_document
date: "2023-11-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Exercises  {#Exercises_DataReshaping}
    
1.  A common task is to take a set of data that has multiple categorical 
    variables and create a table of the number of cases for each combination.
    An introductory statistics textbook contains a dataset summarizing student
    surveys from several sections of an intro class. The two variables of
    interest for us are `Gender` and `Year` which are the students gender and
    year in college.
    a)  Download the dataset and correctly order the `Year` variable using the
        following:
```{r}
Survey <- read.csv('StudentSurvey.csv')
```
    b)  Using some combination of `dplyr` functions, produce a data set with
        eight rows that contains the number of responses for each gender:year
        combination. Make sure your table orders the `Year` variable in the
        correct order of `First Year`, `Sophmore`, `Junior`, and then `Senior`.
        *You might want to look at the following functions: `dplyr::count` and*
        *`dplyr::drop_na`.* 
```{r}
Survey <- Survey %>% 
  drop_na() %>% # remove NA values
  count(Year, Sex) %>% # all gender & year combinations
  arrange(match(Year, c('FirstYear','Sophomore', 'Junior', 'Senior')), .by_group=FALSE)
Survey              # arrange years in order 
```
    c)  Using `tidyr` commands, produce a table of the number of responses in
        the following form:
        
        |   Gender    |  First Year  |  Sophmore  |  Junior   |  Senior   |
        |:-----------:|:------------:|:----------:|:---------:|:---------:|
        |  **Female** |              |            |           |           |  
        |  **Male**   |              |            |           |           | 
```{r}
Survey2 <- Survey %>% 
  pivot_wider( names_from=Year, values_from=n ) # make years columns
Survey2
```
2.  From the book website, there is a .csv file of the daily maximum 
temperature in Flagstaff at the Pulliam Airport. The direction link is at:
[https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv](https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv)
    a)  Create a line graph that gives the daily maximum temperature for 2005.
        *Make sure the x-axis is a date and covers the whole year.*
```{r}
# load temp data
TempData <- read.csv('https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv') 

PrepData <- TempData %>% 
  filter(Year == 2005) %>% # data in 2005
  select(-X) %>%  # remove 'X' col out of preference
  pivot_longer(3:33, names_to = 'Day',values_to = 'MaxTemp') %>% # X1:X31 cols
  mutate(Date=make_date(year=Year, month=Month, day=str_replace_all(Day, pattern='X', replacement='')))
# make date col

PrepData <- drop_na(PrepData) # remove NA values

ggplot(PrepData, aes(x=Date,y=MaxTemp)) + geom_line() + # date vs temp
    labs(title = "Daily Maximum Temperature in Flagstaff (2005)",
       x = "Date",
       y = "Maximum Temperature (°F)")
```

    b)  Create a line graph that gives the monthly average maximum temperature
        for 2013 - 2015. *Again the x-axis should be the date and the axis*
        *spans 3 years.*
```{r}
TempData <- read.csv('https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv') 

MA.PrepData <- TempData %>% 
  filter(Year >= 2013 & Year <= 2015) %>% # years 2013-2015
  select(-X) %>%  
  pivot_longer(3:33, names_to = 'Day',values_to = 'MaxTemp') %>% 
  mutate(Date=make_date(year=Year, month=Month, day=str_replace_all(Day, pattern='X', replacement=''))) 

MA.PrepData <- drop_na(MA.PrepData)

MA <- MA.PrepData %>%
  group_by(Year, Month) %>% # each month, year combination
  summarise(MA.Temp = mean(MaxTemp)) # mean of temp
                      
ggplot(MA, aes(x=as.Date(paste(Year,Month,"01",sep="-")),y=MA.Temp)) +
  geom_line() + # date vs mean monthly temps
  labs(title = "Monthly Average Maximum Temperature in Flagstaff (2005)",
       x = "Date",
       y = "Monthly Average Maximum Temperature (°F)")
```
4. For this problem we will consider two simple data sets.
```{r}
A <- tribble(
  ~Name, ~Car,
  'Alice', 'Ford F150',
  'Bob',   'Tesla Model III',
  'Charlie', 'VW Bug')
    
B <- tribble(
  ~First.Name, ~Pet,
  'Bob',  'Cat',
  'Charlie', 'Dog',
  'Alice', 'Rabbit')
```
    a)  Squish the data frames together to generate a data set with three rows
        and three columns. Do two ways: first using `cbind` and then using one
        of the `dplyr` `join` commands.
```{r}
B <- B %>% arrange(First.Name) # reorder 'First.Name'
B

ABC <- cbind(A,B$Pet) # A + Pet col in B
ABC

ABC2 <- inner_join(A, B, by=c('Name'='First.Name')) # join A & B
ABC2
```
    b)  It turns out that Alice also has a pet guinea pig. Add another row to
        the `B` data set. Do this using either the base function `rbind`, or
        either of the `dplyr` functions `add_row` or `bind_rows`.
```{r}
C <- tibble( First.Name='Alice', Pet='Guinea Pig') # new row
B2 <- add_row(B, C) # add row
B2
```
    c)  Squish the `A` and `B` data sets together to generate a data set with
        four rows and three columns. Do this two ways: first using `cbind` and
        then using one of the `dplyr` `join` commands. Which was easier to
        program? Which is more likely to have an error.
        The inner join is easier to program and by using cbind the program will
        cause an errror because of the differing number of rows in each data 
        set.
        *Error in data.frame(..., check.names = FALSE) :* 
          *arguments imply differing number of rows: 3, 4*
```{r}
#fourthree <- cbind(A, B2$Pet)
#fourthree
```

```{r}
# combine data frames 
four3 <- inner_join(A, B2, by = c("Name" = "First.Name"))
four3
```
5.  Data table joins are extremely common because effective database design
    almost always involves having multiple tables for different types of 
    objects. To illustrate both the table joins and the usefulness of multiple 
    tables we will develop a set of data frames that will represent a credit 
    card company's customer data base. We will have tables for Customers, 
    Retailers, Cards, and Transactions.  Below is code that will create and 
    populate these tables.
```{r}
Customers <- tribble(
  ~PersonID, ~Name, ~Street, ~City, ~State,
  1, 'Derek Sonderegger',  '231 River Run', 'Flagstaff', 'AZ',
  2, 'Aubrey Sonderegger', '231 River Run', 'Flagstaff', 'AZ',
  3, 'Robert Buscaglia', '754 Forest Heights', 'Flagstaff', 'AZ',
  4, 'Roy St Laurent', '845 Elk View', 'Flagstaff', 'AZ')
Customers    
Retailers <- tribble(
  ~RetailID, ~Name, ~Street, ~City, ~State,
  1, 'Kickstand Kafe', '719 N Humphreys St', 'Flagstaff', 'AZ',
  2, 'MartAnnes', '112 E Route 66', 'Flagstaff', 'AZ',
  3, 'REI', '323 S Windsor Ln', 'Flagstaff', 'AZ' )
Retailers
Cards <- tribble(
  ~CardID, ~PersonID, ~Issue_DateTime, ~Exp_DateTime,
  '9876768717278723',  1,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
  '5628927579821287',  2,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
  '7295825498122734',  3,  '2019-9-28 0:00:00', '2022-9-28 0:00:00',
  '8723768965231926',  4,  '2019-9-30 0:00:00', '2022-9-30 0:00:00' ) 
    
Transactions <- tribble(
  ~CardID, ~RetailID, ~DateTime, ~Amount,
  '9876768717278723', 1, '2019-10-1 8:31:23',    5.68,
  '7295825498122734', 2, '2019-10-1 12:45:45',  25.67,
  '9876768717278723', 1, '2019-10-2 8:26:31',    5.68,
  '9876768717278723', 1, '2019-10-2 8:30:09',    9.23,
  '5628927579821287', 3, '2019-10-5 18:58:57',  68.54,
  '7295825498122734', 2, '2019-10-5 12:39:26',  31.84,
  '8723768965231926', 2, '2019-10-10 19:02:20', 42.83) 
    
Cards <- Cards %>% 
  mutate( Issue_DateTime = lubridate::ymd_hms(Issue_DateTime),
          Exp_DateTime   = lubridate::ymd_hms(Exp_DateTime) )
Cards
Transactions <- Transactions %>% 
  mutate( DateTime = lubridate::ymd_hms(DateTime))
Transactions
```
    a)  Create a table that gives the credit card statement for Derek. It 
    should give all the transactions, the amounts, and the store name. Write 
    your code as if the only initial information you have is the customer's 
    name.
   *Hint: Do a bunch of table joins, and then filter for the desired customer*
  *name. To be efficient, do the filtering first and then do the table joins.*
```{r}
CustName <- 'Derek Sonderegger' # given customer name

Customer <- Customers %>% # filter 'Customers' for 'Derek Sonderegger'
  filter( Name == CustName )

Statement <- Customer %>% # join tables by shared cols
  left_join(Cards,by ='PersonID') %>% # Customer$PersonID<->Cards$PersonID
  left_join(Transactions,by='CardID') %>% # Cards$CardID<->Transactions$CardID
  left_join(Retailers,by='RetailID')#Transactions$RetailID<->Retailers$RetailID
Statement

Statement2 <- Statement %>% 
  select(Name.x, Name.y, DateTime, Amount) # show name, store, time, amount
Statement2
```
    b)  Aubrey has lost her credit card on Oct 15, 2019. Close her credit card 
    at 4:28:21 PM and issue her a new credit card in the `Cards` table.
 *Hint: Using the Aubrey's name, get necessary CardID and PersonID and save*
  *those as `cardID` and `personID`. Then update the `Cards` table row that*
  *corresponds to the `cardID` so that the expiration date is set to the time*
  *that the card is closed. Then insert a new row with the `personID` for*
        *Aubrey and a new `CardID` number that you make up.* 
```{r}
CustName2 <- 'Aubrey Sonderegger' # given name

Customer2 <- Customers %>% # filter 'Customers' for 'Aubrey Sonderegger'
  filter( Name == CustName2 )

Info <- Customer2 %>% # use PersonID to get CardID
  left_join(Cards,by ='PersonID') # Customer$PersonID<->Cards$PersonID

cardID <- Info$CardID # save vars
personID <- Info$PersonID
CloseTime <- ymd_hms("2019-10-15 4:28:21 PM") # close card time

Cards <- Cards %>% 
  filter(CardID == cardID) %>% # if card ID is Audrey's, close it
  mutate(Exp_DateTime = CloseTime) # new expiration is close time

NewCardRow <- tibble( # create new card info
  CardID = '3875649120094678',
  PersonID = personID,
  Issue_DateTime = CloseTime, # card issued as old one is closed
  Exp_DateTime = ymd_hms('2022-9-20 0:00:00')
  )

Cards <- bind_rows(Cards, NewCardRow) # issue new card
Cards
```
    c)  Aubrey is using her new card at Kickstand Kafe on Oct 16, 2019 at
        2:30:21 PM for coffee with a charge of $4.98. Generate a new 
        transaction for this action. 
        *Hint: create temporary variables `card`,`retailid`,`datetime`, and*
        *`amount` that contain the information for this transaction and then*
    *write your code to use those. This way in the next question you can just*
    *use the same code but modify the temporary variables. Alternatively, you*
  *could write a function that takes in these four values and manipulates the*
  *tables in the GLOBAL environment using the `<<-` command to assign a result*
   *to a variable defined in the global environment. The reason this is OK is*
    *that in a real situation, these data would be stored in a database and we*
        *would expect the function to update that database.*
```{r}
card <- '3875649120094678' # Audrey's card number
retailid <- 1 # kickstand
datetime <- ymd_hms('2019-10-16 14:30:21')
amount <- 4.98

NewRow <- tibble(
  CardID = card,
  RetailID = retailid,
  DateTime = datetime,
  Amount = amount
)

Transactions <- bind_rows(Transactions, NewRow) # new transaction entry
Transactions
```
    d)  On Oct 17, 2019, some nefarious person is trying to use her OLD credit
        card at REI. Make sure your code in part (c) first checks to see if the
        credit card is active before creating a new transaction. Using the same
        code, verify that the nefarious transaction at REI is denied.
        *Hint: your check ought to look something like this:*
```{r}
card <- '5628927579821287' # Audrey's old card number
retailid <- 3 # REI
datetime <- ymd_hms('2019-10-17 14:30:21')
amount <- 4.98
        
# If the card is currently valid, this should return exactly 1 row.
Valid_Cards <- Cards %>% 
  filter(CardID == card, Issue_DateTime <= datetime, datetime <= Exp_DateTime)
        
# If the transaction is valid, insert the transaction into the table
if( nrow(Valid_Cards) == 1){
    NewRow <- tibble(
    CardID = card,
    RetailID = retailid,
    DateTime = datetime,
    Amount = amount
  )

  Transactions <- bind_rows(Transactions, NewRow)
}else{
  print('Card Denied')
}
```
    e)  Generate a table that gives the credit card statement for Aubrey. It
        should give all the transactions, amounts, and retailer name for both
        credit cards she had during this period.
```{r}
statement <- Customer2 %>% # join tables by shared cols
  left_join(Cards,by ='PersonID') %>% # Customer$PersonID<->Cards$PersonID
  left_join(Transactions,by='CardID') %>% # Cards$CardID<->Transactions$CardID
  left_join(Retailers,by='RetailID')#Transactions$RetailID<->Retailers$RetailID
statement

 statement2 <- statement %>% 
   select(Name.x, CardID, Name.y, DateTime, Amount)
 statement2 # show name, card number, retailer, amount
```
        