---
title: "445-4"
output: pdf_document
date: "2023-10-26"
---

```{r setup, include=FALSE}
library(tidyverse)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises  {#Exercises_DataReshaping}
    
1.  A common task is to take a set of data that has multiple categorical 
    variables and create a table of the number of cases for each combination.
    An introductory statistics textbook contains a dataset summarizing student
    surveys from several sections of an intro class. The two variables of
    interest for us are `Gender` and `Year` which are the students gender and
    year in college.
    a)  Download the dataset and correctly order the `Year` variable using the
        following:
```{r, eval=FALSE}
Survey <- read.csv('http://www.lock5stat.com/datasets/StudentSurvey.csv', na.strings=c('',' ')) 
```
    b)  Using some combination of `dplyr` functions, produce a data set with
        eight rows that contains the number of responses for each gender:year
        combination. Make sure your table orders the `Year` variable in the
        correct order of `First Year`, `Sophmore`, `Junior`, and then `Senior`.
        *You might want to look at the following functions: `dplyr::count` and*
        *`dplyr::drop_na`.* 
```{r}        
 Gender.Year %>% filter(StudentSurvey, Year:Sex)  

```        
    c)  Using `tidyr` commands, produce a table of the number of responses in
        the following form:
        
        |   Gender    |  First Year  |  Sophmore  |  Junior   |  Senior   |
        |:-----------:|:------------:|:----------:|:---------:|:---------:|
        |  **Female** |              |            |           |           |  
        |  **Male**   |              |            |           |           | 
    
2.  From the book website, there is a .csv file of the daily maximum 
temperature in Flagstaff at the Pulliam Airport. The direction link is at: [https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv](https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv)
    a)  Create a line graph that gives the daily maximum temperature for 2005.
        *Make sure the x-axis is a date and covers the whole year.*
    b)  Create a line graph that gives the monthly average maximum temperature
        for 2013 - 2015. *Again the x-axis should be the date and the axis*
        *spans 3 years.*

4.  Data table joins are extremely common because effective database design
    almost always involves having multiple tables for different types of
    objects.To illustrate both the table joins and the usefulness of multiple 
    tables we will develop a set of data frames that will represent a credit 
    card company's customer data base. We will have tables for Customers, 
    Retailers, Cards, and Transactions.  Below is code that will create and 
    populate these tables.
```{r}
Customers <- tribble(
  ~PersonID, ~Name, ~Street, ~City, ~State,
  1, 'Derek Sonderegger',  '231 River Run', 'Flagstaff', 'AZ',
  2, 'Aubrey Sonderegger', '231 River Run', 'Flagstaff', 'AZ',
  3, 'Robert Buscaglia', '754 Forest Heights', 'Flagstaff', 'AZ',
  4, 'Roy St Laurent', '845 Elk View', 'Flagstaff', 'AZ')
    
Retailers <- tribble(
  ~RetailID, ~Name, ~Street, ~City, ~State,
  1, 'Kickstand Kafe', '719 N Humphreys St', 'Flagstaff', 'AZ',
  2, 'MartAnnes', '112 E Route 66', 'Flagstaff', 'AZ',
  3, 'REI', '323 S Windsor Ln', 'Flagstaff', 'AZ' )
    
Cards <- tribble(
  ~CardID, ~PersonID, ~Issue_DateTime, ~Exp_DateTime,
  '9876768717278723',  1,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
  '5628927579821287',  2,  '2019-9-20 0:00:00', '2022-9-20 0:00:00',
  '7295825498122734',  3,  '2019-9-28 0:00:00', '2022-9-28 0:00:00',
  '8723768965231926',  4,  '2019-9-30 0:00:00', '2022-9-30 0:00:00' ) 
    
Transactions <- tribble(
  ~CardID, ~RetailID, ~DateTime, ~Amount,
  '9876768717278723', 1, '2019-10-1 8:31:23',    5.68,
  '7295825498122734', 2, '2019-10-1 12:45:45',  25.67,
  '9876768717278723', 1, '2019-10-2 8:26:31',    5.68,
  '9876768717278723', 1, '2019-10-2 8:30:09',    9.23,
  '5628927579821287', 3, '2019-10-5 18:58:57',  68.54,
  '7295825498122734', 2, '2019-10-5 12:39:26',  31.84,
  '8723768965231926', 2, '2019-10-10 19:02:20', 42.83) 
    
Cards <- Cards %>% 
  mutate( Issue_DateTime = lubridate::ymd_hms(Issue_DateTime),
          Exp_DateTime   = lubridate::ymd_hms(Exp_DateTime) )
Transactions <- Transactions %>% 
  mutate( DateTime = lubridate::ymd_hms(DateTime))
```
    a)  Create a table that gives the credit card statement for Derek. It 
    should give all the transactions, the amounts, and the store name. Write 
    your code as if the only initial information you have is the customer's 
    name.
    *Hint: Do a bunch of table joins, and then filter for the desired customer*
    *name.To be efficient, do the filtering first and then do the table joins.*
    b)  Aubrey has lost her credit card on Oct 15, 2019. Close her credit card 
    at 4:28:21 PM and issue her a new credit card in the `Cards` table.
    *Hint: Using the Aubrey's name, get necessary CardID and PersonID and save*
    *those as `cardID` and `personID`. Then update the `Cards` table row that*
    *corresponds to the `cardID` so that the expiration date is set to thetime*
    *that the card is closed. Then insert a new row with the `personID` for*
        *Aubrey and a new `CardID` number that you make up.* 
    c)  Aubrey is using her new card at Kickstand Kafe on Oct 16, 2019 at
        2:30:21 PM for coffee with a charge of $4.98. Generate a new 
        transaction for this action. 
        *Hint: create temporary variables `card`,`retailid`,`datetime`, and*
        *`amount` that contain the information for this transaction and then*
        *write your code to use those. This way in the next question you can just*
        *use the same code but modify the temporary variables. Alternatively, you*
        *could write a function that takes in these four values and manipulates the*
        *tables in the GLOBAL environment using the `<<-` command to assign a result*
        *to a variable defined in the global environment. The reason this is OK is*
        *that in a real situation, these data would be stored in a database and we*
        *would expect the function to update that database.*
    d)  On Oct 17, 2019, some nefarious person is trying to use her OLD credit
        card at REI. Make sure your code in part (c) first checks to see if the
        credit card is active before creating a new transaction. Using the same
        code, verify that the nefarious transaction at REI is denied.
        *Hint: your check ought to look something like this:*
```{r, eval=FALSE}
card <- '9876768717278723'
retailid <- 2
datetime <- ymd_hms('2019-10-16 14:30:21')
amount <- 4.98
        
# If the card is currently valid, this should return exactly 1 row.
Valid_Cards <- Cards %>% 
  filter(CardID == card, Issue_DateTime <= datetime, datetime <= Exp_DateTime)
        
# If the transaction is valid, insert the transaction into the table
if( nrow(Valid_Cards) == 1){
# Some code to insert the transaction
}else{
  print('Card Denied')
}
```
    e)  Generate a table that gives the credit card statement for Aubrey. It
        should give all the transactions, amounts, and retailer name for both
        credit cards she had during this period.
  
5.  The package `nycflights13` contains information about all the flights that
    arrived in or left from New York City in 2013. This package contains five
    data tables, but there are three data tables we will work with. The data
    table `flights` gives information about a particular flight, `airports`
    gives information about a particular airport, and `airlines` gives
    information about each airline. Create a table of all the flights on
    February 14th by Virgin America that has columns for the carrier,
    destination, departure time, and flight duration. Join this table with
    the airports information for the destination. Notice that because the
    column for the destination airport code doesn't match up between `flights`
    and `airports`, you'll have to use the `by=c("TableA.Col"="TableB.Col")`
    argument where you insert the correct names for `TableA.Col` and 
    `TableB.Col`.







