---
title: "445-14"
output: pdf_document
date: "2023-11-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(latex2exp)
library(plotly)
```

## Exercises  {#Exercises_Advanced_Graphing}


1.  The `infmort` data set from the package `faraway` gives the infant mortality
    rate for a variety of countries. The information is relatively out of date
    (from 1970s?), but will be fun to graph. Visualize the data using by 
    creating scatter plots of mortality vs income while faceting using `region` 
    and setting color by `oil` export status. Utilize a $\log_{10}$ 
    transformation for both `mortality` and `income` axes. This can be done 
    either by doing the transformation inside the `aes()` command or by 
    utilizing the `scale_x_log10()` or `scale_y_log10()` layers. The critical 
    difference is if the scales are on the original vs log transformed scale. 
    Experiment with both and see which you prefer.
    a)  The `rownames()` of the table gives the country names and you should
    create a new column that contains the country names. *`rownames`
```{r}
data(infmort, package='faraway')
rownames <- rownames(infmort) # new vector with country names
infmort <- infmort %>% 
  cbind(rownames) # make 'rownames' a new col
head(infmort)
```
    b)  Create scatter plots with the `log10()` transformation inside the 
    `aes()` command.
```{r}
infmort <- infmort %>%
  drop_na() # remove NA values
ggplot(infmort, aes(x=mortality, y=log10(income))) + 
  geom_point()     # mortality vs income with log10
```
    c)  Create the scatter plots using the `scale_x_log10()` and 
    `scale_y_log10()`.
    Set the major and minor breaks to be useful and aesthetically pleasing.
    Comment on which version you find easier to read.
      I found scaling the y-axis with log10 to be more helpful because the 
      y-axis deals with a much larger range of numbers.
```{r}
ggplot(infmort, aes(x=mortality, y=income)) +
  geom_point() +
  scale_x_log10(breaks=10^(0:3), # major breaks from 0 -> 1,000
                minor  = outer(seq(0,10,by=1), 10^(0:3)) # 9 minor breaks
               ) +                                       # in each major break
               xlab('Mortality (100s of years)')
ggplot(infmort, aes(x=mortality, y=income)) +
  geom_point() +
  scale_y_log10(breaks=10^(0:4), # major breaks from 0 -> 10,000
                minor  = outer(seq(0,10,by=1), 10^(0:4)) # 9 minor breaks
               ) +                                       # in each major break
               ylab('Income (1000s of dollars)')
```
    d)  The package `ggrepel` contains functions `geom_text_repel()` and 
    `geom_label_repel()` that mimic the basic `geom_text()` and `geom_label()`
    functions in `ggplot2`, but work to make sure the labels don't overlap.
    Select 10-15 countries to label and do so using the `geom_text_repel()`
    function.
```{r}
ggplot(infmort, aes(x=mortality, y=income)) +
  geom_point() +
  scale_y_log10(breaks=10^(0:4), 
                minor  = outer(seq(0,10,by=1), 10^(0:4))) + 
  ylab('Income (1000s of dollars)') +
  ggrepel::geom_text_repel(data = infmort,aes(label = rownames))
```

2. Using the `datasets::trees` data, complete the following:
    a)  Create a regression model for $y=$ `Volume` as a function of $x=$ 
    `Height`.
```{r}
head(datasets::trees)
model <- lm( Volume ~ Height, data=trees) # fit regression model
model
```
    b)  Using the `summary` command, get the y-intercept and slope of the
    regression line.
```{r}
summary(model) # summary of model
```
    c)  Using `ggplot2`, create a scatter plot of Volume vs Height. 
```{r}
ggplot(data=trees, aes(y=Volume, x=Height)) + geom_point() # vol vs height
```
    d)  Create a nice white filled rectangle to add text information to using 
    by adding the following annotation layer.
```{r}
ggplot(data=trees, aes(y=Volume, x=Height)) + geom_point() + 
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74, # add rectangle
          fill='white', color='black') 
```
    e)  Add some annotation text to write the equation of the line
        $\hat{y}_i = -87.12 + 1.54 * x_i$ in the text area.
```{r}
ggplot(data=trees, aes(y=Volume, x=Height)) + geom_point() + 
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
           fill='white', color='black') + 
  annotate('text',x=70, y=70, # equation of line
           label = latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$'))
```
    f)  Add annotation to add $R^2 = 0.358$
```{r}
ggplot(data=trees, aes(y=Volume, x=Height)) + geom_point() + 
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
           fill='white', color='black') + 
  annotate('text',x=70, y=70, 
           label = latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$')) +
  annotate('text',x=70, y=65,
           label = latex2exp::TeX('$R^2 = 0.358$')) # r^2 label
```
    g)  Add the regression line in red. The most convenient layer function to 
    uses is `geom_abline()`. It appears that the `annotate` doesn't work with 
    `geom_abline()` so you'll have to call it directly.
```{r}
ggplot(data=trees, aes(y=Volume, x=Height)) + geom_point() + 
  annotate('rect', xmin=65, xmax=75, ymin=60, ymax=74,
           fill='white', color='black') + 
  annotate('text',x=70, y=70, 
           label = latex2exp::TeX('$\\hat{y}_i = -87.12 + 1.54 * x_i$')) +
  annotate('text',x=70, y=65,
           label = latex2exp::TeX('$R^2 = 0.358$')) + # regression line in red
  ggplot2::geom_abline(intercept = -87.12, slope = 1.54, color='red')
```

