---
title: "Assignment4"
author: "Ella Buxton"
date: "2023-09-14"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1

1.  The dataset `ChickWeight` tracks the weights of 48 baby chickens
(chicks) feed four different diets. *Feel free to complete all parts
of the exercise in a*
    *single R pipeline at the end of the problem.*
    a.  Load the dataset using  
```{r}
data(ChickWeight)
```

    b.  Look at the help files for the description of the columns.

```{r}
# ?ChickWeight
```

    c)  Remove all the observations except for observations from day
    10 or day 20. The tough part in this instruction is 
    distinguishing between "and" and "or". Obviously there are no 
    observations that occur from both day 10 AND day 20. Google 'R 
    logical operators' to get an introduction to those, but the 
    short answer is that and is `&` and or is `|`.

```{r}
ChickWeight %>% 
  filter( Time==10 | Time==20 ) # observations from day 10 & 20
head(ChickWeight)
```

    d)  Calculate the mean and standard deviation of the chick 
    weights for each diet group on days 10 and 20.

```{r}
ChickWeight %>% 
  filter( Time==10 | Time==20 ) %>%  # day 10 & 20
  group_by(Diet, Time) %>%  
  summarise(mean.weight=mean(weight), std.weight=sd(weight))
```

2.  The OpenIntro textbook on statistics includes a data set on body dimensions. 
    *Instead of creating an R chunk for each step of this problem, create a*
    *single R pipeline that performs each of the following tasks.*
    a)  Load the file using 
    
```{r}
Body <- read.csv('http://www.openintro.org/stat/data/bdims.csv') # load file
```

    b)  The column `sex` is coded as a 1 if the individual is male and 0 if 
    female. This is a non-intuitive labeling system. Create a new column 
    `sex.MF` that uses labels Male and Female. Use this column for the rest of
    the problem. _Hint: The `ifelse()` command will be__very convenient here. 
    It functions similarly to the same command in Excel._
        
```{r}
Body <- Body %>%
  mutate( Sex.MF = if_else(sex == '1', 'Male', 'Female' ) )
head(Body) # re-label male and female in sex column
```

    c)  The columns `wgt` and `hgt` measure weight and height in kilograms and 
    centimeters (respectively). Use these to calculate the Body Mass Index 
    (BMI) for each individual where 
    $$BMI=\frac{Weight\,(kg)}{\left[Height\,(m)\right]^{2}}$$
        
```{r}
Body <- Body %>% 
  mutate( BMI=wgt/(hgt/100)^2 ) # calculate BMI using wgt & hgt
head(Body)
```

    d)  Double check that your calculated BMI column is correct by examining 
    the summary statistics of the column (e.g. `summary(Body)`). BMI values 
    should be between 18 to 40 or so.  Did you make an error in your 
    calculation? 
    
    I initially made an error in my calculations because I multiplied the 
    height by 100 instead of dividing by 100 to convert the hgt to meters.
    
```{r}
summary(Body$BMI)
```

    e)  The function `cut` takes a vector of continuous numerical data and 
    creates a factor based on your given cut-points. 
```{r}
# Define a continuous vector to convert to a factor
x <- 1:10

# divide range of x into three groups of equal length
cut(x, breaks=3)

# divide x into four groups, where I specify all 5 break points 
cut(x, breaks = c(0, 2.5, 5.0, 7.5, 10))
# (0,2.5] (2.5,5] means 2.5 is included in first group
# right=FALSE changes this to make 2.5 included in the second  
        
# divide x into 3 groups, but give them a nicer
# set of group names
cut(x, breaks=3, labels=c('Low','Medium','High'))
```
Create a new column of in the data frame that divides the age into decades 
(10-19, 20-29, 30-39, etc). Notice the oldest person in the study is 67.
```{r}
Body <- Body %>% 
  mutate( Age.Grp = cut(age, breaks=c(10,20,30,40,50,60,70), right=FALSE))
head(Body) # new column of ages by decade
```

    f)  Find the average BMI for each `Sex.MF` by `Age.Grp` combination.

```{r}
Body <- Body %>% 
  group_by(Sex.MF, Age.Grp) %>%  # group by Sex.MF:Age.Grp
  summarise(Avg.BMI = mean(BMI))  # average BMI for each group
Body                             
```






