---
title: "Assignment5"
output: pdf_document
date: "2023-09-21"
---

```{r setup, include=FALSE}
library(tidyverse)
data('phbirths', package='faraway')
knitr::opts_chunk$set(echo = TRUE)
```

2.  The data set `phbirths` from the `faraway` package contains information on 
    birth weight, gestational length, and smoking status of mother. We'll fit a
    quadratic model to predict infant birth weight using the gestational time.
    
    a)  Create two scatter plots of gestational length and birth weight, one 
    for each smoking status.

```{r}
ggplot(phbirths, aes(x=gestate, y=grams)) + geom_point() + # gestate vs grams
  facet_grid(cols = vars(smoke)) # graph for each smoke status
```

    b)  Remove all the observations that are premature (less than 36 weeks). 
    For the remainder of the problem, only use these full-term babies.

```{r}
phbirths <- phbirths %>% 
  filter(gestate > 36) #  mature observations only
ggplot(phbirths, aes(x=gestate, y=grams)) + geom_point() +
  facet_grid(cols = vars(smoke)) 
```

    c)  Fit the quadratic model 
    
```{r}
model <- lm(grams ~ poly(gestate,2) * smoke, data=phbirths)
#summary(model)
```

    d)  Add the model fitted values to the `phbirths` data frame along with the
    regression model confidence intervals.
        
```{r}
# remove any previous model prediction values that I've added,
# and then add the model predictions 
phbirths <- phbirths %>% 
  dplyr::select( -matches('fit'), -matches('lwr'), -matches('upr') ) %>%
  cbind(predict(model, interval='confidence'))
head(phbirths)
```

    e)  On your two scatter plots from part (a), add layers for the model fits 
    and ribbon of uncertainty for the model fits.

```{r}
ggplot(phbirths, aes(x=gestate, y=grams)) + geom_point() + 
  facet_grid(cols = vars(smoke)) +
  geom_line( aes(y=fit) ) + # model fit
  geom_ribbon( aes( ymin=lwr, ymax=upr), alpha=.3 ) # ribbon of uncertainty
```

    f)  Create a column for the residuals in the `phbirths` data set using any 
    of the following:
    
```{r}
phbirths <- phbirths %>% mutate( residuals = resid(model) ) # residuals column
head(phbirths)
```

2.  The $Uniform\left(a,b\right)$ distribution is defined on x $\in [a,b]$ and 
    represents a random variable that takes on any value of between `a` and `b`
    with equal probability. Technically since there are an infinite number of 
    values between `a` and `b`, each value has a probability of 0 of being 
    selected and I should say each interval of width $d$ has equal probability.
    It has the density function 
    $$f\left(x\right)=\begin{cases}
    \frac{1}{b-a} & \;\;\;\;a\le x\le b\\
    0 & \;\;\;\;\textrm{otherwise}
    \end{cases}$$

    The R function `dunif()` evaluates this density function for the above 
    defined values of x, a, and b. Somewhere in that function, there is a chunk
    of code that evaluates the density for arbitrary values of $x$. Run this 
    code a few times and notice sometimes the result is $0$ and sometimes it is
    $1/(10-4)=0.16666667$.
    
```{r}
a <- 4      # The min and max values we will use for this example
b <- 10     # Could be anything, but we need to pick something
    
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
    
# what is value of f(x) at the randomly selected x value?  
dunif(x, a, b)
```
    
    We will write a sequence of statements that utilizes if statements to 
    appropriately calculate the density of `x`, assuming that `a`, `b` , and 
    `x` are given to you, but your code won't know if `x` is between `a` and 
    `b`. That is, your code needs to figure out if it is and give either 
    `1/(b-a)` or `0`.
    
    a. We could write a set of `if else` statements.
    
```{r, eval=FALSE}
a <- 4
b <- 10
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
        
if( x < a ){
  result <- 0 # value = 0 when less than 'a'
}else if( x <= b ){
  result <- 1/(b-a) # value = 1/(b-a) when greater than a & less than or equal
}else{              # to b
  result <- 0. # value = 0 when greater than 'a' & 'b'
}
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```
        Replace the `????` with the appropriate value, either 0 or 
        $1/\left(b-a\right)$. Run the code repeatedly until you are certain 
        that it is calculating the correct density value.
        
    b.  We could perform the logical comparison all in one comparison. Recall 
        that we can use `&` to mean “and” and `|` to mean “or”. In the 
        following two code chunks, replace the `???` with either `&` or `|` to 
        make the appropriate result.
        
i. 
```{r, eval=FALSE}
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
if( (a<=x) & (x<=b) ){ # if a<=x<=b, value = 1/(b-a)
  result <- 1/(b-a)
}else{
  result <- 0 # if not a<=x<=b, value = 0
}
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```
ii. 
```{r, eval=FALSE}
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
if( (x<a) | (b<x) ){ # if x = less than a OR greater than b, value = 1/(b-a)
  result <- 0
}else{
  result <- 1/(b-a)
}
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```
iii.
```{r, eval=FALSE}
x <- runif(n=1, 0,10)  # one random value between 0 and 10 
result <- ifelse( a<=x & x<=b, 1/(b-a), 0 )
print(paste('x=',round(x,digits=3), '  result=', round(result,digits=3)))
```

3.  I often want to repeat some section of code some number of times. For 
example, I might want to create a bunch plots that compare the density of a 
t-distribution with specified degrees of freedom to a standard normal 
distribution. 
    
```{r, fig.height=3, message=FALSE, warning=FALSE}
library(ggplot2)
df <- 5
N <- 1000
x.grid <- seq(-3, 3, length=N)
data <- data.frame( 
  x = c(x.grid, x.grid),
  y = c(dnorm(x.grid), dt(x.grid, df)),
  type = c( rep('Normal',N), rep('T',N) ) )
    
# make a nice graph
myplot <- ggplot(data, aes(x=x, y=y, color=type, linetype=type)) +
  geom_line() +
  labs(title = paste('Std Normal vs t with', df, 'degrees of freedom'))
    
# actually print the nice graph we made
print(myplot) 
```
    
    a)  Use a `for` loop to create similar graphs for degrees of freedom 
        $2,3,4,\dots,29,30$. 
        
```{r}
library(ggplot2)
df <- 30
N <- 1000
result <- 0

result <- result
for( i in 2:df ){ # loop for df 2:30
  x.grid <- seq(-3, 3, length=N)
data <- data.frame( 
  x = c(x.grid, x.grid),
  y = c(dnorm(x.grid), dt(x.grid, df)),
  type = c( rep('Normal',N), rep('T',N) ) )
 # plot data
myplot <- ggplot(data, aes(x=x, y=y, color=type, linetype=type)) +
  geom_line() +
  labs(title = paste('Std Normal vs t with', i, 'degrees of freedom'))
  result[i] <- result
print(myplot) # print graphs 2:30
}    
```
    
    b)  In retrospect, perhaps we didn't need to produce all of those. Rewrite 
        your loop so that we only produce graphs for 
        $\left\{ 2,3,4,5,10,15,20,25,30\right\}$ degrees of freedom. 
        *Hint: you can just modify the vector in the `for` statement to include*
        *the desired degrees of freedom.*

```{r}
library(ggplot2)
N <- 1000
vector <- c(2,3,4,5,10,15,20,25,30)

result <- result
for( i in vector ){ # loop for df 2:30
  x.grid <- seq(-3, 3, length=N)
data <- data.frame( 
  x = c(x.grid, x.grid),
  y = c(dnorm(x.grid), dt(x.grid, df)),
  type = c( rep('Normal',N), rep('T',N) ) )
 # plot data
myplot <- ggplot(data, aes(x=x, y=y, color=type, linetype=type)) +
  geom_line() +
  labs(title = paste('Std Normal vs t with', i, 'degrees of freedom'))
  result[i] <- result
print(myplot) # print graphs 2:30
}    
```



