---
title: "Assignment1-445"
output: pdf_document
date: "2023-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
```

## Chapter 8 Exercises  {#Exercises_DataStructures}
1. Create a vector of three elements (2,4,6) and name that vector `vec_a`. 
Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors
together and name the result `vec_c`.

```{r}
vec_a <- c(2,4,6)
vec_b <- c(8,10,12)
vec_c <- vec_a + vec_b # adding vec_a & vec_b
vec_c
```

2. Create a vector, named `vec_d`, that contains only two elements (14,20). Add
this vector to `vec_a`. What is the result and what do you think R did (look up
the recycling rule using Google)? What is the warning message that R gives you?
The result added the first and second elements from each vector. But for the 
third element in 'vec_d' it added the third element to the first element in 
'vec_a' & "recycled" it.

Warning: longer object length is not a multiple of shorter object 
length[1] 16 24 20

```{r}
vec_d <- c(14,20)
vec_ad <- vec_a + vec_d # adding vec_a to vec_d
vec_ad
```

3. Next add 5 to the vector vec_a. What is the result and what did R do? Why 
doesn't in give you a warning message similar to what you saw in the previous 
problem?

result is '[1]  7  9 11'. R added 5 to each element in the 'vec_a'. R doesn't 
give you a warning message similar to what I saw in the previous problem 
because 5 is not a vector.

```{r}
vec_a + 5 # adding 5 to each element in vec_a
```

4. Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two 
different ways. 
    a) First using the `seq()` function 
    
```{r}
seq(1,5) # vector from 1 to 5
```
    
    b) Using the `a:b` shortcut.

```{r}
1:5 # also vector from 1 to 5
```

5. Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 
    a) Using the seq() function and 
    
```{r}
seq(2, 20, by=2) # vector from 2 to 20 by 2's
```
    
    b) Using the a:b shortcut and some subsequent algebra. 
    *Hint: Generate the vector 1-10 and then multiple it by 2*.

```{r}
2*(1:10) # also a vector from 2 to 20 by 2's
```

6. Generate a vector of 21 elements that are evenly placed between 0 and 1 
using the `seq()` command and name this vector `x`. 

```{r}
x <- seq(0, 1, length.out=21) # 21 numbers with values from 0 to 1 evenly
                              # placed
x
```

7. Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ 
  using the `rep()` command to replicate the vector c(2,4,8).

```{r}
rep(c(2, 4, 8), 3) # vector repeated 3 times
```

8. Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
  using the `rep()` command. You might need to check the help file for rep() to
  see all of the options that rep() will accept. In particular, look at the 
  optional argument `each=`.
  
```{r}
# ?rep
rep(c(2, 4, 8), each = 4) # repeat each element 4 times
```

10. In this problem, we will work with the matrix 

    \[ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]\]
    
    a) Create the matrix in two ways and save the resulting matrix as `M`.
        i. Create the matrix using some combination of the `seq()` and 
        `matrix()` commands.
        
```{r}
M <- matrix( seq(2, 30, by=2), nrow=3, ncol=5) 
M # 2 to 30 in intervals of 2 in a 3x5 matrix
```
        
        ii. Create the same matrix by some combination of multiple `seq()` 
        commands and either the `rbind()` or `cbind()` command. 
        
```{r}
row_a <- seq(2, 10, by=2)
row_b <- seq(12, 20, by=2)
row_c <- seq(22, 30, by=2)
M2 <- rbind(row_a, row_b, row_c) # bind all rows to create matrix
M2
```

    b) Extract the second row out of `M`.
    
```{r}
M[-(6:10)] # extract second row
```
    
    c) Extract the element in the third row and second column of `M`.

```{r}
M[-c(12)] # extract element in row 3, col 2
```

12. The following code creates a `data.frame` and then has two different 
methods for removing the rows with `NA` values in the column `Grade`. Explain 
the difference between the two.

The first line calls for everything to be included in the output except for 
'NA', and the second line calls for the opposite of 'NA' to be called... so 
everything else will be in the output.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                 Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]
df[  which( !is.na(df$Grade) ), ]
```

14. Create and manipulate a list.
    a) Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131
        
```{r}
x <- c(4,5,6,7,8,9,10)
y <- c(34,35,41,40,45,47,51)
slope <- '2.82'
p.value <- '0.000131'
my.test <-  list(x=x, y=y, slope=slope, p.value=p.value) # create list
str(my.test) # show structure of object
```
        
    b) Extract the second element in the list.
    
```{r}
my.test[[2]] # extract second element
```

    c) Extract the element named `p.value` from the list.

```{r}
my.test[['p.value']] # extract p.value
```

## Chapter 9 Exercises  {#Exercises_DataImport}

1.  Download from GitHub the data file 
    [Example_5.xls](https://github.com/dereksonderegger/444/raw/master/data-raw/Example_5.xls). 
    Open it in Excel and figure out which sheet of data we should import into 
    R. At the same time figure out how many initial rows need to be skipped. 
    Import the data set into a data frame and show the structure of the 
    imported data using the `str()` command. Make sure that your data has 
    $n=31$ observations and the three columns are appropriately named. If you 
    make any modifications to the data file, comment on those modifications.

```{r}
# setwd("/Users/ellabuxton/Desktop/STA444")
data.5 <- read_excel( # import data set
  'Example_5.xls', # data set
  sheet='RawData', # selected sheet
  range='A5:C36') # selected range of data on sheet
str(data.5) # display structure of selected data
```

2.  Download from GitHub the data file 
    [Example_3.xls](https://github.com/dereksonderegger/444/raw/master/data-raw/Example_3.xls). 
    Import the data set into a data frame and show the structure of the 
    imported data using the `tail()` command which shows the last few rows of a
    data table. Make sure the Tesla values are `NA` where appropriate and that 
    both `-9999` and `NA` are imported as NA values. If you make any 
    modifications to the data file, comment on those modifications.

```{r}
data.3 <- read_excel( # import data set
  'Example_3.xls', # data set
  sheet='data', # selected sheet
  range='A1:L34', # selected range of data on sheet
  na=c('NA', -9999)) # show 'NA' instead of values in vector
tail(data.3) # display tail of data
```








